---
title: "Puff_analysis"
author: "nbergum"
date: '2023-04-20'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lme4)
library(readxl)
library(performance)
library(svglite)
library(pbkrtest)
library(lmerTest)
library(car)
library(rstatix)
library(pwr)
library(EnvStats)
```

IV curve visualization and analysis
```{r}
IV_curve <- read.csv("C:\\Users\\nikbe\\Documents\\Vigh Lab\\Puff\\IV_curve.csv")
str(IV_curve)
SumStatsIV <- dplyr::summarise(group_by(IV_curve, treatment, voltage,), 
n = n(),
mean_i = mean(current),
sd_i = sd(current),
se_i = sd_i/sqrt(n))
SumStatsIV

plot_IV <- ggplot(data= SumStatsIV, aes(x=voltage, y=mean_i, color=treatment)) +  geom_point(stat="identity", position=position_dodge(), size=3) + geom_line(aes(group=treatment), linetype="dashed") + geom_errorbar(data= SumStatsIV, aes(ymin=mean_i-se_i, ymax=mean_i+se_i), width=.25)  + scale_x_continuous(limits = c(-100,20)) + scale_color_manual(values=c("black", "red")) +  ylab("Current Amplitude (pA)") +  xlab("Voltage (mV)") + geom_hline(yintercept=0, alpha=.4) + geom_vline(xintercept = 0, alpha=.4)
plot_IV

ggsave(file="plot_IV.svg", plot=plot_IV, width=6, height=4)

IV_curve2 <- droplevels(IV_curve[!IV_curve$cell == 'cell g',])
IV_curve2
SumStatsIV2 <- dplyr::summarise(group_by(IV_curve2, treatment, voltage,), 
n = n(),
mean_i = mean(current),
sd_i = sd(current),
se_i = sd_i/sqrt(n))
SumStatsIV2

plot_IV2 <- ggplot(data= SumStatsIV2, aes(x=voltage, y=mean_i, color=treatment)) +  geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(aes(group=treatment)) + geom_errorbar(data= SumStatsIV2, aes(ymin=mean_i-se_i, ymax=mean_i+se_i), width=.25) + geom_hline(yintercept=0, linetype="dotted") +  scale_x_continuous(limits = c(-100,20)) + geom_vline(xintercept = 0, linetype="dotted")
plot_IV2

IV_curve$voltage <- as.factor(IV_curve$voltage)
lm_IV <- lmer(croot ~ treatment*voltage+(1|cell), data= IV_curve)
plot(lm_IV, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_IV)
shapiro.test(resid(lm_IV))
anova(lm_IV)
emm_IV1  <- emmeans::emmeans(lm_IV, pairwise ~ treatment|voltage)
emm_IV1
```


Effect of morphine on GABA-puff IPSC: summary, visualization and analysis
```{r}
morph_puff <- read.csv("C:\\Users\\nikbe\\Documents\\Vigh Lab\\Puff\\morphine_puff.csv")

SumStatsMOR <- dplyr::summarise(group_by(morph_puff, treatment, ), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n)
)
SumStatsMOR

plot_IampMOR <- ggplot() + geom_bar(data= SumStatsMOR, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=morph_puff, aes(x= treatment, y=amplitude), size=2) + geom_line(data=morph_puff, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsMOR, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + xlab("Treatment") + ylim(0,850) + scale_fill_manual(values=c("grey","red","skyblue")) + theme_bw()
plot_IampMOR

ggsave(file="plot_IampMOR.svg", plot=plot_IampMOR, width=6, height=4)

morph_puff2 <- subset(morph_puff, washout == "yes")
morph_puff2

SumStatsMOR2 <- dplyr::summarise(group_by(morph_puff2, treatment, ), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n)
)
SumStatsMOR2

plot_IampMOR2 <- ggplot() + geom_bar(data= SumStatsMOR2, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=morph_puff2, aes(x= treatment, y=amplitude), size=2) + geom_line(data=morph_puff2, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsMOR2, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + xlab("Treatment") + ylim(0,850) + scale_fill_manual(values=c("grey","red","skyblue")) + theme_bw()
plot_IampMOR2

#ggsave(file="plot_IampMOR2.svg", plot=plot_IampMOR2, width=6, height=4)

lm_Ipeakmor <- lmer(amplitude ~ treatment + (1|cell), data = morph_puff)
plot(lm_Ipeakmor, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_Ipeakmor)
shapiro.test(resid(lm_Ipeakmor))
anova(lm_Ipeakmor)
emmeans::emmeans(lm_Ipeakmor, pairwise ~ treatment)

#lm_Ipeakmor2 <- lmer(amplitude ~ treatment + (1|cell), data = morph_puff2)
#plot(lm_Ipeakmor2, type=c("p","smooth"), col.line=1)
#lattice::qqmath(lm_Ipeakmor2)
#shapiro.test(resid(lm_Ipeakmor2))
#anova(lm_Ipeakmor2)
#emmeans::emmeans(lm_Ipeakmor2, pairwise ~ treatment)
```

Effect of DAMGO on GABA-puff IPSC: summary, visualization and analysis
```{r}
damgo_puff <- read.csv("C:\\Users\\nikbe\\Documents\\Vigh Lab\\Puff\\damgo_puff.csv")

SumStatsDAMGO <- dplyr::summarise(group_by(damgo_puff, treatment, ), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n)
)
SumStatsDAMGO

plot_IampDAMGO <- ggplot() + geom_bar(data= SumStatsDAMGO, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=damgo_puff, aes(x= treatment, y=amplitude), size=2) + geom_line(data=damgo_puff, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsDAMGO, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + xlab("Treatment") + ylim(0,850) + scale_fill_manual(values=c("grey","red","blue")) + theme_bw()
plot_IampDAMGO

ggsave(file="plot_IampDAMGO.svg", plot=plot_IampDAMGO, width=6, height=4)

damgo_puff2 <- subset(damgo_puff, washout == "YES")
damgo_puff2

SumStatsDAMGO2 <- dplyr::summarise(group_by(damgo_puff2, treatment, ), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n)
)

SumStatsDAMGO2

SumStatsDAMGO2 <- ggplot() + geom_bar(data= SumStatsDAMGO2, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=damgo_puff2, aes(x= treatment, y=amplitude), size=2) + geom_line(data=damgo_puff2, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsDAMGO2, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + xlab("Treatment") + ylim(0,850) + scale_fill_manual(values=c("grey","red","blue")) + theme_bw()
SumStatsDAMGO2

#ggsave(file="SumStatsDAMGO2.svg", plot=SumStatsDAMGO2, width=6, height=4)

lm_Ipeakdamgo <- lmer(amplitude ~ treatment + (1|cell), data = damgo_puff)
plot(lm_Ipeakdamgo, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_Ipeakdamgo)
shapiro.test(resid(lm_Ipeakdamgo))
anova(lm_Ipeakdamgo)
emmeans::emmeans(lm_Ipeakdamgo, pairwise ~ treatment)

#lm_Ipeakdamgo2 <- lmer(amplitude ~ treatment + (1|cell), data = damgo_puff2)
#plot(lm_Ipeakdamgo2, type=c("p","smooth"), col.line=1)
#lattice::qqmath(lm_Ipeakdamgo2)
#shapiro.test(resid(lm_Ipeakdamgo2))
#anova(lm_Ipeakdamgo2)
#emmeans::emmeans(lm_Ipeakdamgo2, pairwise ~ treatment)
```

```{r}
temp3 <- morph_puff %>% filter(treatment == "control" | treatment == "washout") %>%
  group_by(cell) %>%
  summarise(baseline = mean(amplitude))
temp4 <- morph_puff %>% filter(treatment == "morphine") %>%
  group_by(cell) %>%
  summarise(effect = amplitude)

percent_mor <- merge(temp3, temp4, all = T)

percent_mor$perc_inc <- (percent_mor$effect - percent_mor$baseline)/percent_mor$baseline
percent_mor$trt <- 'morphine' 
percent_mor

percent_damgo <- damgo_puff %>% group_by(cell) %>% dplyr::summarise(base = (amplitude[treatment=="control"] + amplitude[treatment=="NLX"])/2, perc_inc = (amplitude[treatment=="DAMGO"]/ base)-1)

temp5 <- damgo_puff %>% filter(treatment == "control" | treatment == "NLX") %>%
  group_by(cell) %>%
  summarise(baseline = mean(amplitude))
temp6 <- damgo_puff %>% filter(treatment == "DAMGO") %>%
  group_by(cell) %>%
  summarise(effect = amplitude)

percent_damgo <- merge(temp5, temp6, all = T)

percent_damgo$perc_inc <- (percent_damgo$effect - percent_damgo$baseline)/percent_damgo$baseline

percent_damgo$trt <- 'DAMGO' 
percent_damgo

percent_mu <- merge(percent_damgo, percent_mor, all = T)
percent_mu

SumStatsMOR1 <- dplyr::summarise(group_by(percent_mu, trt), n = n(), mean_perc = mean(perc_inc),sd_perc = sd(perc_inc), se_perc = sd_perc/sqrt(n),
)
SumStatsMOR1

plot_incMOR1 <- ggplot(data= SumStatsMOR1, aes(x=trt, y=mean_perc, fill=trt)) + geom_bar(stat="identity", width = 0.5) + geom_point(data=percent_mu, aes(y=perc_inc), size=2) + geom_errorbar(data= SumStatsMOR1, aes(ymin=mean_perc-se_perc, ymax=mean_perc+se_perc), width=.25) + scale_y_continuous(labels = scales::percent)
plot_incMOR1
```
Establishing baseline and standard deviation for non-responsiveness in IGABA
```{r}
damgo_puff2
morph_puff2

percent.damgo <- damgo_puff2 %>% group_by(cell) %>% dplyr::summarise(base = (amplitude[treatment=="control"] + amplitude[treatment=="NLX"])/2, perc_inc = (amplitude[treatment=="DAMGO"]/ base)-1)

percent.morph <- morph_puff2 %>% group_by(cell) %>% dplyr::summarise(base = (amplitude[treatment=="control"] + amplitude[treatment=="washout"])/2, perc_inc = (amplitude[treatment=="morphine"]/ base)-1)

percent.damgo
percent.morph

percent.mu_opioid <- merge(percent.damgo, percent.morph, all = T)
percent.mu_opioid

percent.mu_opioid$percent <- percent.mu_opioid$perc_inc*100
percent.mu_opioid <- data.frame(percent.mu_opioid)

boxplot(percent.mu_opioid$percent)

test_mu <- rosnerTest(percent.mu_opioid$percent, k = 2)
test_mu

percent.mu_opioid$scale_percent <- scale(percent.mu_opioid$percent)
hist(percent.mu_opioid$scale_percent, breaks = 25)
summary(percent.mu_opioid$scale_percent)

percent.mu.opioid <- subset(percent.mu_opioid, percent < 15)
percent.mu.opioid


SumStatsMuR <- dplyr::summarise(group_by(percent.mu.opioid,), 
        n = n(), 
        mean_perc = mean(percent), 
        sd_perc = sd(percent), 
        se_perc = sd_perc/sqrt(n), 
        )
SumStatsMuR
```


Dopamine responsiveness of ipRGCs: summary and visualization
```{r}
da_response <- read.csv("C:\\Users\\nikbe\\Documents\\Vigh Lab\\Puff\\da_responsive.csv")
str(da_response)
SumStatsDA <- dplyr::summarise(group_by(da_response, treatment, DA_responsive), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n),
mean_auc = mean(AUC),
sd_auc = sd(AUC),
se_auc = sd_auc/sqrt(n)
)
SumStatsDA

da_response$treatment <- as.factor(da_response$treatment)
da_response$DA_responsive <- as.factor(da_response$DA_responsive)
da_response$cell <- as.factor(da_response$cell)

plot_Iamp <- ggplot() + geom_bar(data= SumStatsDA, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=da_response, aes(x= treatment, y=amplitude), size=2) + geom_line(data=da_response, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsDA, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + 
  xlab("Treatment") + scale_fill_manual(values=c("grey","red","blue")) + theme_bw() + ylim(0,850) + facet_grid(.~ DA_responsive)
plot_Iamp

ggsave(file="plot_Iamp.pdf", plot=plot_Iamp, width=6, height=4)
ggsave(file="plot_Iamp.svg", plot=plot_Iamp, width=6, height=4)

plot_I_auc <- ggplot() + geom_bar(data= SumStatsDA, aes(treatment, mean_auc, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=da_response, aes(x= treatment, y=AUC), size=2) + geom_line(data=da_response, aes(x=treatment, y=AUC, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsDA, aes(x=treatment, ymin=mean_auc-se_auc, ymax=mean_auc+se_auc, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Charge Transfer (pQ)") + xlab("Treatment") + scale_fill_manual(values=c("grey","red","blue")) + theme_bw() + ylim(0,1500) + facet_grid(.~ DA_responsive)
plot_I_auc

ggsave(file="plot_Iamp.pdf", plot=plot_Iamp, width=6, height=4)
ggsave(file="plot_Iamp.svg", plot=plot_Iamp, width=6, height=4)
ggsave(file="plot_I_auc.svg", plot=plot_I_auc, width=6, height=4)

lm_IpeakDA <- lmer(amplitude ~ treatment*DA_responsive + (1|cell), data = da_response)
lm_IpeakDA
#check_model(lm_IpeakDA)
plot(lm_IpeakDA, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_IpeakDA)
shapiro.test(resid(lm_IpeakDA))
anova(lm_IpeakDA)
emm  <- emmeans::emmeans(lm_IpeakDA, pairwise ~ treatment|DA_responsive)
emm

p <- power.anova.test(groups = 3, 
between.var = 89.45, within.var = 24.23, 
power=0.9,sig.level=0.05)
p

lm_IAUCDA <- lmer(AUC ~ treatment*DA_responsive + (1|cell), data = da_response)
plot(lm_IAUCDA, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_IAUCDA)
shapiro.test(resid(lm_IAUCDA))
anova(lm_IAUCDA)
emmeans::emmeans(lm_IAUCDA, pairwise ~ treatment|DA_responsive)
```

Calculating the percent responsiveness of a cells to dopamine: visualization and statistics
```{r}
da_response
percent_da <- da_response %>% group_by(cell, DA_responsive) %>% dplyr::summarise(base = (amplitude[treatment=="control"] + amplitude[treatment=="SCH39166"])/2, perc_inc = (amplitude[treatment=="dopamine"]/ base)-1)
percent_da$trt <- 'DA' 
percent_da

SumStatsDopamine <- dplyr::summarise(group_by(percent_da, DA_responsive), 
        n = n(), 
        mean_perc = mean(perc_inc), 
        sd_perc = sd(perc_inc), 
        se_perc = sd_perc/sqrt(n), 
        )
SumStatsDopamine

plot_da_perc <-  ggplot() + geom_bar(data= SumStatsDopamine, aes(DA_responsive, mean_perc, fill= DA_responsive), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=percent_da, aes(x= DA_responsive, y=perc_inc), size=2) + geom_errorbar(data= SumStatsDopamine, aes(x=DA_responsive, ymin=mean_perc-se_perc, ymax=mean_perc+se_perc, fill = DA_responsive), position=position_dodge(width=0.9), width=.25) + scale_y_continuous(labels = scales::percent, limits = c(-0.1,0.8)) + theme_bw() + ylab("Percent Peak Current Change")
plot_da_perc

ggsave(file="plot_da_perc.pdf", plot=plot_da_perc, width=6, height=4)
ggsave(file="plot_da_perc.svg", plot=plot_da_perc, width=6, height=4)

res.ftest <- var.test(perc_inc ~ DA_responsive, data = percent_da)
res.ftest

lm_da <- lm(perc_inc ~ DA_responsive, data = percent_da)

shapiro.test(rstandard(lm_da))
qqPlot(percent_da$perc_inc)

res <- wilcox.test(perc_inc ~ DA_responsive, data = percent_da)
res
```
SKF383939 responsiveness of ipRGCs: summary and visualization
```{r}
skf_response <- read.csv("C:\\Users\\nikbe\\Documents\\Vigh Lab\\Puff\\skf_sch_response.csv")
skf_response$treatment <- factor(skf_response$treatment, levels = c("control", "SKF38393", "SCH39166"))
str(skf_response)
SumStatsSKF <- dplyr::summarise(group_by(skf_response, treatment, DA_responsive), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n),
)
SumStatsSKF

skf_response$treatment <- as.factor(skf_response$treatment)
skf_response$DA_responsive <- as.factor(skf_response$DA_responsive)
skf_response$cell <- as.factor(skf_response$cell)


plot_IampSKF <- ggplot() + geom_bar(data= SumStatsSKF, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=skf_response, aes(x= treatment, y=amplitude), size=2) + geom_line(data=skf_response, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsSKF, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + scale_fill_manual(values=c("grey","red","blue")) + xlab("Treatment") + theme_bw() + ylim(0,850) + facet_grid(.~ DA_responsive)
plot_IampSKF

ggsave(file="plot_IampSKF.pdf", plot=plot_IampSKF, width=6, height=4)
ggsave(file="plot_IampSKF.svg", plot=plot_IampSKF, width=6, height=4)

lm_IpeakSKF <- lmer(amplitude ~ treatment*DA_responsive + (1|cell), data = skf_response)
lm_IpeakSKF
#check_model(lm_IpeakDA)
plot(lm_IpeakSKF, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_IpeakSKF)
shapiro.test(resid(lm_IpeakSKF))
anova(lm_IpeakSKF)
emmSKF  <- emmeans::emmeans(lm_IpeakSKF, pairwise ~ treatment|DA_responsive)
emmSKF
```

Calculating the percent responsiveness of a cells to SKF383939: visualization and statistics
```{r}
temp1 <- skf_response %>% filter(treatment == "control" | treatment == "SCH39166") %>% group_by(cell, DA_responsive) %>% summarise(baseline = mean(amplitude))
temp2 <- skf_response %>% filter(treatment == "SKF38393") %>%group_by(cell, DA_responsive) %>% summarise(effect = amplitude)
percent_skf <- merge(temp1, temp2, all = T)
percent_skf$perc_inc <- (percent_skf$effect - percent_skf$baseline)/percent_skf$baseline
  
#percent_skf <- skf_response %>% group_by(cell, DA_responsive) %>% dplyr::summarise(perc_inc = (amplitude[treatment=="SKF38393"]/ amplitude[treatment=="SKF38393"])-1)
percent_skf$trt <- 'SKF' 
percent_skf
  
SumStatsSKF1 <- dplyr::summarise(group_by(percent_skf, trt, DA_responsive), 
        n = n(), 
        mean_perc = mean(perc_inc), 
        sd_perc = sd(perc_inc), 
        se_perc = sd_perc/sqrt(n), 
        )
SumStatsSKF1

plot_skf_perc <-  ggplot() + geom_bar(data= SumStatsSKF1, aes(DA_responsive, mean_perc, fill= DA_responsive), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=percent_skf, aes(x= DA_responsive, y=perc_inc), size=2) + geom_errorbar(data= SumStatsSKF1, aes(x=DA_responsive, ymin=mean_perc-se_perc, ymax=mean_perc+se_perc, fill = DA_responsive), position=position_dodge(width=0.9), width=.25) + scale_y_continuous(labels = scales::percent, limits = c(-0.1,0.8)) + theme_bw() + ylab("Percent Peak Current Change")
plot_skf_perc

ggsave(file="plot_skf_perc.pdf", plot=plot_skf_perc, width=6, height=4)
ggsave(file="plot_skf_perc.svg", plot=plot_skf_perc, width=6, height=4)

res.ftest <- var.test(perc_inc ~ DA_responsive, data = percent_skf)
res.ftest

lm_skf <- lm(perc_inc ~ DA_responsive, data = percent_skf)

shapiro.test(rstandard(lm_skf))
qqPlot(percent_skf$perc_inc)

t.test(perc_inc ~ DA_responsive, data = percent_skf, paired = FALSE)

d_value1<- cohens_d(perc_inc ~ DA_responsive, data = percent_skf, paired = FALSE)
d_value1

pwr.t.test(d=-3.338243, power=0.99, sig.level=.05,type="two.sample",alternative="less")
```
Frequency histogram for D1R-responsiveness
```{r}
percent_dopamine <- merge(percent_skf, percent_da, all = T)
percent_dopamine$percent <- percent_dopamine$perc_inc*100
percent_dopamine 
percent_dopamine = percent_dopamine %>% 
  dplyr::select(-c("base", "baseline","effect"))

percent_dopamine

SumStatsDopamineR <- dplyr::summarise(group_by(percent_dopamine, DA_responsive), 
        n = n(), 
        mean_perc = mean(percent), 
        sd_perc = sd(percent), 
        se_perc = sd_perc/sqrt(n), 
        )
SumStatsDopamineR

dopamine.ftest <- var.test(perc_inc ~ DA_responsive, data = percent_dopamine)
dopamine.ftest

lm_dopamine <- lm(percent ~ DA_responsive, data = percent_dopamine)

shapiro.test(rstandard(lm_dopamine))
qqPlot(percent_dopamine$percent)

res_da <- wilcox.test(percent ~ DA_responsive, data = percent_dopamine)
res_da

response_histo1 <- ggplot(percent_dopamine, aes(x=percent, color=DA_responsive, fill= DA_responsive)) + geom_histogram(aes(y=..density..), binwidth=2.5, alpha=0.2, position="identity") + geom_density(alpha=0.6) + geom_vline(data=SumStatsDopamineR, aes(xintercept=mean_perc, color=DA_responsive), linetype="dotted", size=1.5)

response_histo1

ggsave(file="response_histo1.svg", plot=response_histo1, width=6, height=4)

response_histo2 <- ggplot(percent_dopamine, aes(x=percent, color=DA_responsive, fill= DA_responsive)) + geom_histogram(aes(y=..density..), binwidth=1.5, alpha=0.2, position="identity") + geom_density(alpha=0.6) + geom_vline(data=SumStatsDopamineR, aes(xintercept=mean_perc, color=DA_responsive), linetype="dotted", size=2) + geom_vline(xintercept =11, linetype="dashed", size=0.7) 

response_histo2

ggsave(file="response_histo2.svg", plot=response_histo2, width=6, height=4)

response_histo3 <- ggplot(percent_dopamine, aes(x=percent, color=DA_responsive, fill= DA_responsive)) + geom_histogram(aes(y=..density..), binwidth=1.5, alpha=0.2, position="identity") + geom_density(alpha=0.6) + geom_vline(data=SumStatsDopamineR, aes(xintercept=mean_perc, color=DA_responsive), linetype="dotted", size=1.5) + 
  annotate('rect', xmin=-13, xmax=11, ymin=0, ymax=Inf, alpha=.2, fill='red') + geom_vline(xintercept =11, linetype="dashed", size=0.7) + geom_vline(xintercept =-13, linetype="dashed", size=0.7) 

response_histo3

ggsave(file="response_histo3.svg", plot=response_histo3, width=6, height=4)
```

Effect of  50 uM forskolin on GABA-puff IPSC: summary, visualization and analysis
```{r}
forskolin <- read.csv("C:\\Users\\nikbe\\Documents\\Vigh Lab\\Puff\\fsk_puff.csv")
str(forskolin)
percentdelta <- forskolin %>% group_by(cell) %>% dplyr::summarise(perc_inc = ((amplitude[treatment=="FSK"])/ (amplitude[treatment=="control"]))-1)
percentdelta$trt <- 'fsk50uM' 
percentdelta

SumStatsFSK <- dplyr::summarise(group_by(forskolin, treatment, ), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n),
mean_auc = mean(AUC),
sd_auc = sd(AUC),
se_auc = sd_auc/sqrt(n)
)
SumStatsFSK

SumStatsInc <- dplyr::summarise(group_by(percentdelta, trt,), 
n = n(),
mean_inc = mean(perc_inc),
sd_inc = sd(perc_inc),
se_inc = sd_inc/sqrt(n))

SumStatsInc 

forskolin$treatment <- as.factor(forskolin$treatment)
forskolin$cell <- as.factor(forskolin$cell)

plot_Iampf50 <- ggplot() + geom_bar(data= SumStatsFSK, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=forskolin, aes(x= treatment, y=amplitude), size=2) + geom_line(data=forskolin, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsFSK, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + xlab("Treatment") + scale_fill_manual(values=c("grey","red")) + theme_bw() + ylim(0,950)
plot_Iampf50

ggsave(file="plot_Iampf50.pdf", plot=plot_Iampf50, width=6, height=4)
ggsave(file="plot_Iampf50.svg", plot=plot_Iampf50, width=6, height=4)

plot_incFSK <- ggplot(data= SumStatsInc, aes(x=trt, y=mean_inc, fill=trt)) + geom_bar(stat="identity", width = 0.5) + geom_point(data=percentdelta, aes(y=perc_inc), size=2) + geom_errorbar(data= SumStatsInc, aes(ymin=mean_inc-se_inc, ymax=mean_inc+se_inc), width=.25) + scale_y_continuous(labels = scales::percent, limits = c(-0.1,0.4)) 
plot_incFSK

bartlett.test(amplitude ~ treatment, forskolin)
t.test(formula = amplitude ~ treatment, data = forskolin, paired = TRUE)


shapiro.test(percentdelta$perc_inc)
t.test(percentdelta$perc_inc, mu =0, alternative = "greater")
```

Effect of  5 uM forskolin on GABA-puff IPSC: summary, visualization and analysis
```{r}
fsk5um <- read.csv("C:\\Users\\nikbe\\Documents\\Vigh Lab\\Puff\\fsk5uM_response.csv")
str(fsk5um)
percentdelta5 <- fsk5um %>% group_by(cell) %>% dplyr::summarise(perc_inc = ((amplitude[treatment=="FSK"])/ (amplitude[treatment=="control"]))-1)
percentdelta5$trt <- 'fsk5uM' 
percentdelta5

SumStatsFSK5 <- dplyr::summarise(group_by(fsk5um, treatment, ), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n),
)
SumStatsFSK5

SumStatsInc5 <- dplyr::summarise(group_by(percentdelta5, trt,), 
n = n(),
mean_inc = mean(perc_inc),
sd_inc = sd(perc_inc),
se_inc = sd_inc/sqrt(n))
SumStatsInc5

fsk5um$treatment <- as.factor(fsk5um$treatment)
fsk5um$cell <- as.factor(fsk5um$cell)

plot_Iampf5 <- ggplot() + geom_bar(data= SumStatsFSK5, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=fsk5um, aes(x= treatment, y=amplitude), size=2) + geom_line(data=fsk5um, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsFSK5, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + xlab("Treatment") + scale_fill_manual(values=c("grey","red")) + theme_bw() + ylim(0,600)
plot_Iampf5

ggsave(file="plot_Iampf5.pdf", plot=plot_Iampf5, width=6, height=4)
ggsave(file="plot_Iampf5.svg", plot=plot_Iampf5, width=6, height=4)

plot_incFSK5 <- ggplot(data= SumStatsInc5, aes(x=trt, y=mean_inc, fill=trt)) + geom_bar(stat="identity", width = 0.5) + geom_point(data=percentdelta5, aes(y=perc_inc), size=2) + geom_errorbar(data= SumStatsInc5, aes(ymin=mean_inc-se_inc, ymax=mean_inc+se_inc), width=.25) + scale_y_continuous(labels = scales::percent, limits = c(-0.1,0.4))
plot_incFSK5

bartlett.test(amplitude ~ treatment, fsk5um)
t.test(formula = amplitude ~ treatment, data = fsk5um, paired = TRUE)
shapiro.test(percentdelta5$perc_inc)
t.test(percentdelta5$perc_inc, mu =0, alternative = "greater")
```

Effect of  15.8 uM forskolin on GABA-puff IPSC: summary, visualization and analysis
Comparing different concentrations of forskolin
```{r}
fsk15um <- read.csv("C:\\Users\\nikbe\\Documents\\Vigh Lab\\Puff\\fsk15uM_response1.csv")
str(fsk15um)
fsk15um_cs <- filter(fsk15um, internal == "CsGlc")
percentdelta15_cs <- fsk15um_cs %>% group_by(cell) %>% dplyr::summarise(perc_inc = ((amplitude[treatment=="FSK"])/ (amplitude[treatment=="control"]))-1)
percentdelta15_cs$trt <- 'fsk15uM' 
percentdelta15_cs

fsk15um_kt <- filter(fsk15um, internal == "KT5720")
percentdelta15_kt <- fsk15um_kt %>% group_by(cell) %>% dplyr::summarise(perc_inc = ((amplitude[treatment=="FSK"])/ (amplitude[treatment=="control"]))-1)
percentdelta15_kt$trt <- 'fsk15uM+KT5720'
percentdelta15_kt

percentdelta15 <- merge(percentdelta15_cs, percentdelta15_kt, , all = T)
percentdelta15

SumStatsFSK15 <- dplyr::summarise(group_by(fsk15um, treatment, internal, ), 
n = n(),
mean_amp = mean(amplitude),
sd_amp = sd(amplitude),
se_amp = sd_amp/sqrt(n),
)
SumStatsFSK15

plot_Iampf15 <- ggplot() + geom_bar(data= SumStatsFSK15, aes(treatment, mean_amp, fill= treatment), position= "dodge", stat = "identity", width = 0.5) + geom_point(data=fsk15um, aes(x= treatment, y=amplitude), size=2) + geom_line(data=fsk15um, aes(x=treatment, y=amplitude, group=cell), linetype="dotted") + geom_errorbar(data= SumStatsFSK15, aes(x=treatment, ymin=mean_amp-se_amp, ymax=mean_amp+se_amp, fill = treatment), position=position_dodge(width=0.9), width=.25) + ylab("Peak Current Amplitude (pA)") + xlab("Treatment") + scale_fill_manual(values=c("grey","red")) + theme_bw() + facet_grid(.~ internal) + ylim(0,650)
plot_Iampf15

ggsave(file="plot_Iampf15.pdf", plot=plot_Iampf15, width=6, height=4)
ggsave(file="plot_Iampf15.svg", plot=plot_Iampf15, width=6, height=4)

lm_Ipeak_fsk<- lmer(amplitude ~ treatment*internal + (1|cell), data = fsk15um)
lm_Ipeak_fsk
plot(lm_Ipeak_fsk, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_Ipeak_fsk)
shapiro.test(resid(lm_Ipeak_fsk))
anova(lm_Ipeak_fsk)
emm_fsk1  <- emmeans::emmeans(lm_Ipeak_fsk, pairwise ~ treatment|internal)
emm_fsk1
emm_fsk2  <- emmeans::emmeans(lm_Ipeak_fsk, pairwise ~ internal|treatment)
emm_fsk2

SumStatsInc15 <- dplyr::summarise(group_by(percentdelta15, trt,), 
n = n(),
mean_inc = mean(perc_inc),
sd_inc = sd(perc_inc),
se_inc = sd_inc/sqrt(n))
SumStatsInc15

plot_incFSK15 <- ggplot(data= SumStatsInc15, aes(x=trt, y=mean_inc, fill=trt)) + geom_bar(stat="identity", width = 0.5) + geom_point(data=percentdelta15, aes(y=perc_inc), size=2) + geom_errorbar(data= SumStatsInc15, aes(ymin=mean_inc-se_inc, ymax=mean_inc+se_inc), width=.25) + scale_y_continuous(labels = scales::percent, limits = c(-0.1,0.5)) + scale_fill_manual(values=c("red2","purple")) + ylab("Percent Peak Current Change")
plot_incFSK15

ggsave(file="plot_incFSK15.svg", plot=plot_incFSK15, width=6, height=4)

bartlett.test(perc_inc ~ trt, percentdelta15)
t.test(formula = perc_inc ~ trt, data = percentdelta15, paired = FALSE)

d_value<- cohens_d(perc_inc ~ trt, data = percentdelta15, paired = FALSE)
d_value

pwr.t.test(d=3.165836,n=5,sig.level=.05,type="two.sample",alternative="two.sided")

```

```{r}
fsk_a <- merge(percentdelta5, percentdelta, all = T)
fsk_all <- merge(fsk_a, percentdelta15, , all = T)
fsk_all

SumStatsPercFSK <- dplyr::summarise(group_by(fsk_all, trt,), 
n = n(),
mean_inc = mean(perc_inc),
sd_inc = sd(perc_inc),
se_inc = sd_inc/sqrt(n))
SumStatsPercFSK

SumStatsPercFSK

SumStatsPercFSK$trt <- factor(SumStatsPercFSK$trt, levels=c('fsk5uM', 'fsk15uM+KT5720','fsk15uM', 'fsk50uM'))

plot_incFSK_all <- ggplot(data= SumStatsPercFSK, aes(x=trt, y=mean_inc, fill=trt)) + geom_bar(stat="identity", width = 0.5) + geom_point(data=fsk_all, aes(y=perc_inc), size=2) + geom_errorbar(data= SumStatsPercFSK, aes(ymin=mean_inc-se_inc, ymax=mean_inc+se_inc), width=.25) + scale_y_continuous(labels = scales::percent, limits = c(-0.1,0.75)) + ylab("Percent Change in Current Amplitude") + scale_fill_manual(values=c("red", "purple", "red2","red4")) +  ylab("Percent Peak Current Change")
plot_incFSK_all

ggsave(file="plot_incFSK_all.svg", plot=plot_incFSK_all, width=6, height=4)

lm_fsk <- lm(perc_inc ~ trt, data = fsk_all)
plot(lm_fsk)

shapiro.test(rstandard(lm_fsk))
anova(lm_fsk)

em_fsk <- emmeans::emmeans(lm_fsk, pairwise ~ trt)
em_fsk
```